A multi-dimensional extension of the Lightweight Temporal Compression method

### 2018-10-17 paper.tex
Added intro

\author{Bo Li, Tristan Glatard\\
  Department of Computer Science and Software Engineering\\ Concordia University, Montreal, Quebec, Canada\\
  {first.last}@concordia.ca \vspace*{-0.5cm}}

\maketitle

\begin{abstract}
Lightweight Temporal Compression
(LTC~\cite{schoellhammer2004lightweight}) is among the stream
compression method that provides the highest compression rate for the
lowest resource (CPU, RAM) consumption. As such, it makes it a very
good candidate for the compression of data streams acquired in
embedded, low-power systems such as connected
objects, systems on module (SoMs), etc The type of data acquired on
such systems, however, is often multi-dimensional. For instance
accelerometers and gyroscopes usually measure variables along 3
directions, these variables being distinct but related through the
actual movement of the sensor. In this paper, we investigate the
extension of LTC to higher dimensions. First, we provide a formulation of the algorithm
in an arbitrary vectorial space of dimension $n$. Furthermore, we
implement the algorithm for the infinity and Euclidean norms, in spaces of
dimension 2D+t and 3D+t. Finally, we compare the multi-dimensional LTC
to the 1-dimensional version, on acceleration data acquired on a SoM
device.
\end{abstract}

@@

Section~\ref{sec:ltc} provides some background on the LTC algorithm.
Section~\ref{sec:extension} presents our formulation in dimension $n$.

% Some related works from Bo's report

Compression reduces radio usage but increases resource usage.

% transmitted, received points

% contributions: formalization of LTC algorithm, extension to n d, extension to any norm, open-source implementation.

@@

The algorithm \emph{receives} a stream of \emph{data points} $x_i$
at times $t_i$ ($i \in \mathbb{N}$), and it \emph{transmits} a stream of data points $\xi_i$
at times $\tau_i$ ($i \in \mathbb{N}$). To simplify the notations, we assume that:
\begin{equation*}
\forall k \in \mathbb{N}, \  \exists ! i \in \mathbb{N} \  \tau_k = t_i
\end{equation*}
That is, transmission times coincide with reception times.
We define the \emph{centered received points} as follows:
\begin{equation*}
\forall k \in \mathbb{N}\ , \forall j \in \mathbb{N^*},\ (u^k_j, y^k_j) = (t_{i+j}, x_{i+j}), \quad \mathrm{where\ }i\mathrm{\ is\ s.t.}\ t_i = \tau_k.
\end{equation*}
and:
\begin{equation*}
\forall k \in \mathbb{N},\  (u^k_0, y^k_0) = (\tau_k, \xi_k).
\end{equation*}
This definition is such that $y^k_j$ is the $j^{th}$ data point received
after the $k^{th}$ transmission.
Figure~\ref{fig:ltc} illustrates the notations and algorithm.
\begin{figure}
\includegraphics[width=\columnwidth]{ltc.pdf}
\caption{Illustration of the LTC algorithm and notations used. Red dots indicate transmitted points. Dashed lines
represent the high line and the low line at times when a point is transmitted.}
\label{fig:ltc}
\end{figure}
\subsection{LTC algorithm}

@@

\section{Implementation}
Our code is available at ... (make a clean release!)

@@

\section{Experiments and Results}

>>>>

\author{Bo Li$^1$, Omid Sarbishei$^2$, Tristan Glatard$^1$\\
  $^1$ Department of Computer Science and Software Engineering, Concordia University, Montreal, QC, Canada \\
  $^2$ Motsai, Saint Bruno, QC, Canada.}
\maketitle

\begin{abstract}
Lightweight Temporal Compression
(LTC) is among the lossy stream
compression methods that provide the highest compression rate for the
lowest resource (CPU, RAM) consumption. As such, it makes it a very
good candidate for the compression of data streams acquired in
energy-constrained systems such as systems on module (SoMs). The type of data acquired on
such systems, however, is often multi-dimensional. For instance 
accelerometers and gyroscopes usually measure variables along 3 
directions, these variables being distinct but related through the 
actual movement of the sensor. In this paper, we investigate the 
extension of LTC to higher dimensions. First, we provide a formulation 
of the algorithm in an arbitrary vectorial space of dimension $n$. 
Then, we implement the algorithm for the infinity and Euclidean norms, 
in spaces of dimension 2D+t and 3D+t. We evaluate our implementation on 
3D acceleration streams acquired by a SoM during human activities. 
Results show that the 3D implementation of LTC can save up to 20\% in 
energy consumption for low-paced activities such as walking. The 
implementation with the infinity norm only requires about 100~B of 
memory.
\end{abstract}

@@

% Compression of stream data
% need references
Compressing data streams is required in many contexts, to reduce 
storage and network requirements of data analytics. In particular, it is 
highly beneficial to energy-constrained devices such as Systems on 
Modules (SoM), as they reduce the amount of energy-consuming radio transmissions.
However, resource-intensive algorithms such as~\cite{x},~\cite{y} 
or~\cite{z} are not well-suited for energy-constrained systems, due to 
the limited memory available on these systems (typically a few KB), and 
the energy consumption associated with CPU usage. On such systems, 
compression algorithms need to find a trade-off between reducing 
network communications and increasing memory and CPU usage. As 
discussed in~\cite{zordan2014performance}, linear compression methods 
provide a very good compromise between these two factors, leading to 
substantial energy reduction.

% lossy vs lossless compression

% LTC summary
% define transmitted, received points, copmression ratio
The Lightweight Temporal Compression method 
(LTC~\cite{schoellhammer2004lightweight}) has been designed 
specifically for energy-constrained systems, initially sensor networks. 
It approximates data points by a piecewise linear function that 
guarantees an upper bound on the reconstruction error. LTC is a very 
lightweight method, with memory requirements in $\mathcal{O}(1)$. 
However, LTC has only been described for 1D streams, which makes it 
limited for multi-dimensional streams such as the ones coming from 
accelerometers, gyroscopes and other movement sensors.

In this paper, we extend LTC to dimension $n$. To do so, we propose an 
algebraic formulation of the algorithm that also yields a 
norm-independent expression of it. We implement our extension on 
Motsai's Neblina module, and we test it on 3D acceleration streams 
acquired during human exercises, namely biceps curling, walking and 
running. Our implementation of LTC is available as free software on GitHub\todo{add link}.

We assume that the stream consists of a sequence of data points 
received at uneven intervals. The compression algorithm 
\emph{transmits} fewer points than it receives. The transmitted points 
might be part of the stream, or computed from stream elements. The 
\emph{compression ratio} is the ratio between the number of received 
points and the number of transmitted points. An application would 
reconstruct the stream from the transmitted points: the 
\emph{reconstruction error} is the maximum absolute difference between 
a data point of the reconstructed stream, and the corresponding data 
point in the original stream.

% Outline
Section~\ref{sec:ltc} provides some background on the LTC algorithm, 
and formalizes the description initially proposed 
in~\cite{schoellhammer2004lightweight}. Section~\ref{sec:extension} 
presents our norm-independent extension to dimension $n$, and 
Section~\ref{sec:implementation} describes our implementation. 
Section~\ref{sec:results} reports on experiments to validate our 
implementation, and evaluate the impact of n-dimensional LTC on SoM 
energy consumption.
\begin{figure}
\includegraphics[width=\columnwidth]{ltc.pdf}
\caption{Illustration of the LTC algorithm and notations used. Red dots indicate transmitted points. Dashed lines
represent the high line and the low line at times when a point is transmitted.}
\label{fig:ltc}
\end{figure}

@@

The algorithm \emph{receives} a stream of \emph{data points} $x_i$
at times $t_i$ ($i \in \mathbb{N}$), and it \emph{transmits} a stream of data points $\xi_i$
at times $\tau_i$ ($i \in \mathbb{N}$). To simplify the notations, we assume that:
\begin{equation*}
\forall k \in \mathbb{N}, \  \exists ! i \in \mathbb{N} \  \tau_k = t_i
\end{equation*}
That is, transmission times coincide with reception times.
We define the \emph{centered received points} as follows:
\begin{equation*}
\forall k \in \mathbb{N}\ , \forall j \in \mathbb{N^*},\ (u^k_j, y^k_j) = (t_{i+j}, x_{i+j}), \quad \mathrm{where\ }i\mathrm{\ is\ s.t.}\ t_i = \tau_k.
\end{equation*}
and:
\begin{equation*}
\forall k \in \mathbb{N},\  (u^k_0, y^k_0) = (\tau_k, \xi_k).
\end{equation*}
This definition is such that $y^k_j$ is the $j^{th}$ data point received
after the $k^{th}$ transmission.
Figure~\ref{fig:ltc} illustrates the notations and algorithm.

@@

\section{Implementation}
\label{sec:implementation}
Our code is available at ... (make a clean release!)

@@

\section{Experiments and Results}
\label{sec:results}
